Collaborative Visual Analysis on the Web with RCloud

### 2015-03-31 interviews.tex
further tweaking and pruning

By default, all notebooks are publicly visible in RCloud, and notebooks can be
found through the notebook tree or search. However, the method for sharing that
users mentioned most is by sending links through email. Clint says, ``If my supervisor
wants to see what I've done or QA it, I can just send her a link.''

@@

More selective ways to search will be needed as the volume of notebooks
continues to grow. We are considering tagging notebooks filtering the notebook
tree by tag, measuring the quality of notebooks by their forks as well as stars,
and searching within a single user's notebooks.

@@

\subsection{Other workflows}

Although most of the analysts we spoke to use and appreciates the sharing
features, RCloud is less popular as a tool for exploratory data analysis and
development.

Much of this is because of familiarity with other tools. Horace: ``for other
people who use RCloud vs RStudio, they will always think that they want to
prototype on RStudio. Maybe because it's more familiar for them. So, they work
on RStudio, on a small data set, and they fix up everything, make sure
everything is right, and then they place it on RCloud for sharing.''

Many of the analysts we spoke to preferred using a text editor to refine and
store their work, pasting it to an R command line to try it out. Rick describes,
``I have the discipline of having a file that describes what I'm doing with the
commands that I'm using so that I can go back and recreate it or pass it on to
someone else, so it's a little bit like having an RCloud notebook, it's not
necessarily executable but most of the commands that I've typed are in there.''

But the web interface of RCloud doesn't cut it for him. ``I find that the style of editing is clunky. I would never edit like that
if I'm doing ordinary work. I usually use vi.'' ``I mean it's nice to have it
saved, but there's this trade-off between it making it easier for me to present
something or to save something, and my ease of typing and correcting and things
in a plain old editor window.''

Kenny works much the same way, rearranging the file so the good code is at the top of
the file. Then he saves the file once he's got the right code at
the top of the file (and often shares the work by pasting it into an RCloud notebook).

Running anything on the Web can also be a source of frustration.  Poor network
connections can cause web sockets to die, and Corey complains ``What frustrates me is
that if I do leave this window for what feels like 5 or 7 minutes, it
disconnects and I have to reconnect and run the whole notebook again.'' Raif
whines that ``being web-based, it's obviously slower than something you would be
running on your machine.'' For Byers: even the little lag of sending a command
to a server and receiving the result is intolerable

Taylor asks, ``why would I want to use RCloud over my
current setup? If it's just me, I like my text editor and terminal. There's
nothing that I want that those two don't give me.''

@@

\subsection{Proliferation of notebook cells}
RCloud's notebook interface combines editable Markdown with a command line
interface.

The interface takes some learning. Clint says ``it took me a couple of weeks of
looking at it to become comfortable because I saw the cells and it just threw me
for a total loop. I mean, it's a good idea because [...] I can run it all at
once if I want or I can break these into sections for either debugging or
staging purposes. I really like it now, but when I first saw it, [it was] very
confusing [...] I think I prefer it now, because you can cut your code up.''

Just the difference between a cell and the command prompt can confuse and slow
down users. Rick: ``if I was typing into one of the cells near that top, I had
to think `I'm editing this cell and then I have to execute it', and if was doing
the one at the bottom I could type it but then it would automatically execute
[by pressing enter], but then it became a standard cell and I have to edit
it. It was just another set of modes that I didn't like''

@@

Users also felt that cells do not capture the right level of granularity: they
either hold too much code or too little. Rick says that RCloud's cell structure
``tempts me to type a big long thing and then run the whole thing, as opposed to
typing a few little pieces that I'm thinking of on the inner parts and then put
them together to make the big thing'' as he would do on the command line. When
Taylor uses the R command line, he copies and pastes ``like 5-10 lines of code,
so when something breaks, I get an error message on that one line, and I can
up-arrow and change it and fix it, whereas in RCloud I have to run a whole cell,
so the only way to get that same functionality is if every line's in one
cell. Because if it runs lines 1, 2, 3, then 4 fails, I have to either run the
whole cell again, or have line 4 in its own cell. I can't rerun just one line.''

@@

\subsection{A sea of notebooks}
RCloud is starting to be a victim of its own success, as there are now too many
notebooks.  There are over 5000 notebooks on the research instance, of which
more than 500 have been starred and more than 350 have been forked.

Rick doesn't find the notebook tree very useful for this reason: ``I
don't necessarily need to see everybody's notebook that uses RCloud'' ``every
time I do something new, I get a new notebook and so now my notebooks are maybe
50 or 60. That's enough to think about just on my own, but if I've got
everybody's 50 or 60 sitting on my display, I find that it's more than I want to
know about.''

>>>>

By default, all notebooks are publicly visible in RCloud, and notebooks can be
found through the notebook tree or search. However, the method for sharing that
users mentioned most is sending links through email. Clint says, ``If my supervisor
wants to see what I've done or QA it, I can just send her a link.''

@@

More selective ways to search will be needed as the volume of notebooks
continues to grow. We explore some ideas in the Discussion below.

@@

\subsection{Exploring elsewhere}

Although most of the analysts use and appreciate the sharing features,
RCloud is less popular as a tool for exploratory data analysis. Every
analyst with prior R experience still does analysis in another tool and
pastes their code into RCloud for sharing.

This is partly out of familiarity. Taylor asks, ``why would I want to use
RCloud over my current setup? If it's just me, I like my text editor and
terminal. There's nothing that I want that those two don't give me.''

%% Much of this is because of familiarity with other tools. Horace: ``for other
%% people who use RCloud vs RStudio, they will always think that they want to
%% prototype on RStudio. Maybe because it's more familiar for them. So, they work
%% on RStudio, on a small data set, and they fix up everything, make sure
%% everything is right, and then they place it on RCloud for sharing.''

%% Many of the analysts we spoke to preferred using a text editor to refine and
%% store their work, pasting it to an R command line to try it out.

But there are interface reasons as well. Rick describes a common workflow:
\begin{quote}
I have the discipline of having a file that describes what I'm doing, with the
commands that I'm using, so that I can go back and recreate it or pass it on to
someone else. So it's a little bit like having an RCloud notebook. It's not
necessarily executable but most of the commands that I've typed are in there.
\end{quote}

However, the web interface of RCloud doesn't cut it for him:
\begin{quote}
I find that the style of editing is clunky. I would never edit like that if
I'm doing ordinary work. I usually use vi. [...] I mean, it's nice to have
it saved, but there's this trade-off between it making it easier for me to
present something or to save something, and my ease of typing and
correcting and things in a plain old editor window.
\end{quote}

Kenny also works with text file and command line, rearranging a file so the
good code is at the top of the file. Then he saves the file once he's got
the right code at the top, and often shares the work by pasting it into an
RCloud notebook.

%% Running anything on the Web can also be a source of frustration.  Poor network
%% connections can cause web sockets to die, and Corey complains ``What frustrates me is
%% that if I do leave this window for what feels like 5 or 7 minutes, it
%% disconnects and I have to reconnect and run the whole notebook again.'' Raif
%% whines that ``being web-based, it's obviously slower than something you would be
%% running on your machine.'' For Byers, even the little lag of sending a command
%% to a server and receiving the result is intolerable

@@

\subsection{Cells versus the command line}
RCloud's notebook interface combines editable Markdown with a command line
interface.

The interface takes some learning. Clint says
\begin{quote}
It took me a couple of weeks of
looking at it to become comfortable because I saw the cells and it just threw me
for a total loop. I mean, it's a good idea because [...] I can run it all at
once if I want or I can break these into sections for either debugging or
staging purposes. I really like it now, but when I first saw it, [it was] very
confusing [...] I think I prefer it now, because you can cut your code up.
\end{quote}

Just the difference between a cell and the command prompt can confuse and slow
down users. Rick explains:
\begin{quote}
If I was typing into one of the cells near that top, I had
to think ``I'm editing this cell and then I have to execute it,'' and if was doing
the one at the bottom I could type it but then it would automatically execute
[by pressing enter], but then it became a standard cell and I have to edit
it. It was just another set of modes that I didn't like
\end{quote}

@@

Users also felt that cells do not capture the right level of granularity: they
either hold too much code or too little. Rick says that RCloud's cell structure
``tempts me to type a big long thing and then run the whole thing, as opposed to
typing a few little pieces that I'm thinking of on the inner parts and then put
them together to make the big thing'' as he would do on the command line. When
Taylor uses the R command line, he copies and pastes ``like 5-10 lines of code,
so when something breaks, I get an error message on that one line, and I can
up-arrow and change it and fix it, whereas in RCloud I have to run a whole cell,
so the only way to get that same functionality is if every line's in one
cell.''

@@

\subsection{A sea of notebooks}
RCloud is starting to be a victim of its own success, as there are now too many
notebooks.  There are over 5200 notebooks on the research instance, of which
more than 500 have been starred and more than 350 have been forked.

Rick doesn't find the notebook tree very useful for this reason: ``I
don't necessarily need to see everybody's notebook that uses RCloud [...] Every
time I do something new, I get a new notebook and so now my notebooks are maybe
50 or 60. That's enough to think about just on my own, but if I've got
everybody's 50 or 60 sitting on my display, I find that it's more than I want to
know about.''
